---
title: "Task 2"
author: "Hannah Irish"
date: "2023-01-26"
output: html_document
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(AICcmodavg)
```


# **Task 2: California Coast Oxygen Saturation**

The data in this project is from the California Cooperative Oceanic Fisheries Investigations (CalCOFI) 70+ year time series measuring water quality factors off the coast of California including salinity, temperature, oxygen, phosphate and later silicate, nitritie, chlorophyll, etc.


### Read in the data using read_csv()
```{r, warning=FALSE, message=FALSE}
sea_samples <- read_csv(here("data_task2", "calcofi_seawater_samples.csv"))
```

### Create models for multiple linear regression
```{r}
model_1 <- lm(o2sat ~ t_deg_c + salinity + po4u_m, data=sea_samples)
model_2 <- lm(o2sat ~ t_deg_c + salinity + po4u_m + depth_m, data=sea_samples)
```

### Calculate AIC and BIC for models
```{r, output=FALSE}
AIC_1 <- AICc(model_1)
AIC_2 <- AICc(model_2)

BIC_1 <- BIC(model_1)
BIC_2 <- BIC(model_2)

```


### Peform ten-fold cross validaation of the two models using root-mean-square error as the scoring method.

```{r}
folds <- 10 ##set # of folds (#-fold validation)
folds_vector <- rep(1:folds, length_out = nrow(sea_samples)) ##make a vector of 1 to the # of rows you have

set.seed(33) ##to make sure I get the same results each time for debugging

sea_fold <- sea_samples %>%
  mutate(group = sample(folds_vector, size = n(), replace = FALSE))
```








